---
title: "Reproducible research assign 1"
output: html_document
---

Load the activity.csv to R and process the data. 

```{r}
library(data.table)
dt <- fread("activity.csv")
```

Ignore the missing value from the data set and 
calculate the total number of steps taken per day and 
plot this. 

```{r}
date.sum <- dt[,.(steps.sum = sum(steps)),by=.(date)]
hist(date.sum$steps.sum)
mean.steps.total.per.day <- mean(date.sum$steps.sum,na.rm=T)
median.steps.total.per.day <- median(date.sum$steps.sum, na.rm=T)
```

Mean of the total number of steps take per day is `r (mean.steps.total.per.day)`.

Median of the total number of steps take per day is `r (median.steps.total.per.day)`.

Following plot will show a time series plot of the 5-minute inverval (x-axis) and 
the average number of steps taken, averaged across all days (y-axis)

```{r}
# assign 0 to NA
dt[is.na(dt$steps)] <- 0

# group by interval
avg.step <- dt[,.(inter.step.avg = mean(steps)),by=.(interval)]

# 
plot(inter.step.avg ~ interval, avg.step, type="l",
     xlab="5-minute interval",
     ylab="average number of steps taken, average acrooss all days")

sorted_avg <- avg.step[rev(order(inter.step.avg))]
```

5-minute intgerval at **`r sorted_avg$interval[1]`** contains the maximum number of steps. 
```{r}
orgdt <- fread("activity.csv")

# Imputing missing values
tt <- orgdt[is.na(orgdt$steps)]
```

There are **`r length(tt$steps) `** missing values in the dataset.

I am using mean value for the 5-minute interval to replace the missing value for
that 5-minute interval. Following R code will show you how to do that.

```{r}
# Devised a strategy 
# Use the mean 5-min interval to fill in the missing data
# Get NAs from steps variable 

NAs <- is.na(orgdt$steps)

# add inter.step.avg from avg.step to orgdt 
orgdt$avgsteps <- avg.step$inter.step.avg
# Replace orgdt$steps with orgdt$avgsteps is orgdt$steps is NA
orgdt$steps <- as.numeric(orgdt$steps)
orgdt$steps[NAs] <- orgdt$avgsteps[NAs]

rm(NAs)
# Remove the avgsteps column 
orgdt$avgsteps <- NULL
# orgdt <- orgdt[,.(steps,date,interval)]

# After removing the avgsteps column, we have a dataset with no NAs
new.date.sum <- orgdt[,.(steps.sum = sum(steps)),by=.(date)]

# Draw a histogram plot 
hist(new.date.sum$steps.sum)

new.mean.steps.total.per.day <- mean(new.date.sum$steps.sum)
new.median.steps.total.per.day <- median(new.date.sum$steps.sum)
```
Mean of the total number of steps take per day without missing values is **`r (new.mean.steps.total.per.day)`**.

Median of the total number of steps take per day without missing values is **`r (new.median.steps.total.per.day)`**.


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
